Incentivizing Exploration by Heterogeneous Users

### 2018-02-05 algorithm.tex
finished edits up to payment

% move to colt.tex eventually

\newcommand{\e}{\mathrm{e}}
\providecommand{\Kth}[1]{\ensuremath{{#1}^{\rm th}}}

\def\QED{{\phantom{x}} \hfill \ensuremath{\rule{1.3ex}{1.3ex}}}

\newcommand{\extraproof}[1]{\rm \trivlist \item[\hskip \labelsep{\bf Proof of #1. }]}
\def\endextraproof{\QED \endtrivlist}

\def\emptyproof{\rm \trivlist \item[\hskip \labelsep{\bf Proof. }]}
\def\endemptyproof{\endtrivlist}

@@

\dkcomment{I moved the assumptions to an earlier section.
I am saving here some text that was in that subsection.}

@@

The key technical lemma bounds the probability of mistakes in myopic play.
Intuitively, because estimates of the arms' attribute vectors become
exponentially more accurate with increasing phases $s$,
errors from myopic play should decrease exponentially.
That intuition is captured by the following lemma.

@@

\dkcomment{Is it problematic that we use $\sigma$ for two things here?}
Let $\tau$ be any stopping time that is almost surely \dkcomment{What
  exactly do we mean by this here?} in phase $s$,
with respect to the filtration
$\mathcal{F}_{t} = \sigma (\Pull{1}, \ldots, \Pull{t},
                          \PayA{1}, \ldots, \PayA{t},
                          \ObsV{1}, \ldots, \ObsV{t})$.
Then, for all $s \geq \LatePhase$,
\begin{align*}
  \ProbC[\AgV \sim \AgentDist]{%
  \argmax \SET{\AgentV{\tau} \cdot \ArmEV{\tau}{i}} \neq \Best{\AgentV{\tau}}}{%
  \AgentV{\tau} \cdot (\ArmV{\Best{\AgentV{\tau}}} - \ArmV{\Second{\AgentV{\tau}}})
  > 2\Diam d x}
& \leq 24 \ARMNUM d \exp\left(-\frac{1.8 s x^2}{16 \sigma^2} \right).
\end{align*}
\end{lemma}

@@

\begin{extraproof}{Lemma~\ref{lem:round-prob}}
Consider any agent type \AgentV{\tau} at the (random) time $\tau$ such that
$\AgentV{\tau} \cdot (\ArmV{\Best{\AgentV{\tau}}} - \ArmV{\Second{\AgentV{\tau}}})
> 2\Diam d x$.
We will use Lemma~\ref{lem:ACI-inequality} to show that
for any arm $i$ and attribute $j = 1, \ldots, d$,
with probability at least
$1 - 24 \exp \left( \frac{-1.8 s x^2}{16\sigma^2} \right)$,
the estimated attribute $j$ for arm $i$ is close enough to the true attribute: 
$|\ArmE{\tau}{i}{j} - \Arm{i}{j}| \leq x$.
By a union bound over all $i, j$, this bound holds simultaneously for all
arms and coordinates with probability at least
$1 - 24 \ARMNUM d \cdot \exp \left( \frac{-1.8 s x^2}{16\sigma^2} \right)$.

Then, consider any type $\AgV = \AgentV{\tau}$ in the high-probability event,
and let $i \neq \Best{\AgentV{\tau}}$ be any other arm.
By the conditioning of \AgentV{\tau}, the true reward from arm $i$ is
much smaller than from the best arm:
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{i}) > 2\Diam d x$.
Because each coordinate of \ArmEV{\tau}{\Best{\AgV}} and of
\ArmEV{\tau}{i} is estimated accurately to within $x$, 
we get that 
$\AgV \cdot (\ArmEV{\tau}{\Best{\AgV}} - \ArmV{\Best{\AgV}})
\geq - \Diam d x$
and
$\AgV \cdot (\ArmEV{\tau}{i} - \ArmV{i}) \leq \Diam d x$.
Hence,

\begin{align*}
\AgV \cdot (\ArmEV{\tau}{\Best{\AgV}} - \ArmEV{\tau}{i})
& =
\AgV \cdot (\ArmEV{\tau}{\Best{\AgV}} - \ArmV{\Best{\AgV}})
+ \AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{i})
+ \AgV \cdot (\ArmV{i} - \ArmEV{\tau}{i})\\
& > -\Diam d x + 2 \Diam d x - \Diam d x
\; = \; 0,
\end{align*}
which means that the agent with type \AgV will in fact choose
the optimal arm in the high-probability event.

It remains to prove that with high probability, all arms' coordinates
are estimated sufficiently accurately for type \AgentV{\tau}.
Thereto, fix an arm $i$ and attribute $j$.
For any $n \geq 1$, let $k_s$ be the time step right after arm $i$ has
been pulled for the \Kth{n} time.
\dkcomment{I'm writing $n$ instead of $s$ here because $S_s$ would
  look ugly, and $n$ is consistent with the lemma.
  Or we could rename $S$ to something else.}
Define
$S_n := \frac{n \cdot (\ArmE{k_n}{i}{j} - \Arm{i}{j})}{2 \sigma}$
to be the sum of all attribute-$j$ noise components up to and
including the \Kth{n} pull of arm $i$,
\dkcomment{I'm not sure if we have defined the notation for components
  of vectors in the \emph{superscript}. Let's make sure not to forget it.}
renormalized to have standard deviation $1/2$.
\dkcomment{Is this the correct scaling.}
The $S_n$ define an unbiased half-subgaussian random walk,
and we can therefore apply Lemma~\ref{lem:ACI-inequality} to them.
Specifically, we set $b(n) = \frac{n x^2}{16 \sigma^2}$,
and obtain that

@@

Finally, this choice of $J$ ensures that
$\ArmE{k_J}{i}{j} = \ArmE{\tau}{i}{j}$,
and we have thus shown that
$|\ArmE{\tau}{i}{j} - \Arm{i}{j}| \leq x$ for all $i,j$.
\end{extraproof}

@@

Using the above lemmas, we are ready to prove Theorem~\ref{rst:budget}. We include the proof of Theorem~\ref{rst:budget} in the appendix. Below, we bound the expected number of payments for our algorithm.
\begin{lemma}
The expected number of payments for Algorithm~\ref{Alg1} is bounded above by $O(\ARMNUM^2)$.
\label{lemma:numP}
\end{lemma}
\begin{proof}
If $|u_{i}^{j}-u_{i,t}^{j}|\leq \lambda$ is true $\forall i$, $\forall j$, then we know for those $\AgV\in \{\AgV:\AgV\cdot u_{B(\AgV)}-\max_{j\neq B(\AgV)}\{\AgV \cdot u_{j}\}> 2mW\lambda\}$, they will correctly identify their best arm. Thus we know, in the $n^{th}$ round, if $|u_{i}^{j}-u_{i,t}^{j}|\leq \frac{p^{-1}(\frac{p}{2})}{2 \Diam d}$ $\forall i$ and $\forall j$, and $n^{-1}\leq p/2$, we do not need to incentivize any arms. In order to have $n^{-1}\leq \frac{p}{2}$, we need $n\geq \frac{2}{p}$. Denote $n_1=\max\{n_{0}, \frac{2}{p}\}$. Denote $\delta_{0}=p^{-1}(\frac{p}{2})>0$ (because of Assumption~\ref{A1}).

Define $\tau_{n}^{i}$ to be the first time we pull arm $i$ in the $n^{th}$ round. Then
\begin{align}
\sum_{t=1}^{\infty}\mathbbm{1}\{c(t)>0\} =\sum_{n=1}^{\infty}\sum_{i=1}^{N}\mathbbm{1}\{c(\tau_{n}^{i})>0\}. \nonumber
\end{align}

The cumulative expected number of payments is bounded above by:
\begin{align}
&E\left[\sum_{t=1}^{\infty}\mathbbm{1}\{c(t)>0\}\right] \nonumber \\
=&\sum_{n=1}^{\infty}\sum_{i=1}^{N}P(c(\tau_{n}^{i})>0) \nonumber \\
\leq &\sum_{n=n_{1}}^{\infty}\sum_{i=1}^{N}P\left(\exists i,j:|u_{i}^{j}-u_{i,\tau_{n}^{i}}^{j}|>\frac{p^{-1}(\frac{p}{2})}{2 \Diam d}\right)+\sum_{n=1}^{n_1}\ARMNUM \nonumber \\
\leq &\sum_{n=n_{1}}^{\infty}\sum_{i=1}^{N}24 \ARMNUM d
       \exp\left(\frac{-1.8 n\delta_{0}^2}{64 \Diam^2 d^2\sigma^2}\right) +\sum_{n=1}^{n_1}\ARMNUM \nonumber \\
\leq & \sum_{n=n_{1}}^{\infty}24Nm \exp\left(\frac{-1.8 n\delta_{0}^2}{64\Diam^2 d^2\sigma^2}\right)\times \ARMNUM+\sum_{n=1}^{n_1}\ARMNUM \nonumber  \\
\leq&24 \ARMNUM^2 d \frac{1}{\exp(\frac{1.8\delta_{0}}{64\Diam^2 d^2\sigma^2})-1} + Nn_1, \nonumber
\end{align}

Thus, we know the expected number of payments is bounded above by $O(\ARMNUM^2)$.

\end{proof}

\subsection{Bounding the Total Regret}

>>>>

@@

@@

The key technical lemma captures the intuition that
the estimates of the arms' attribute vectors become
exponentially more accurate with increasing phases $s$.
That intuition is captured by the following lemma.

@@

\dkcomment{Is it problematic that we use $\sigma$ for two things here?}
Let $\tau$ be any stopping time that is almost surely \dkcomment{What
  exactly do we mean by this here?} in phase $s$,
with respect to the filtration
$\mathcal{F}_{t} = \sigma (\Pull{1}, \ldots, \Pull{t},
                          \PayA{1}, \ldots, \PayA{t},
                          \ObsV{1}, \ldots, \ObsV{t})$.
Then, for all phases $s \geq \LatePhase$,
with probability at least 
$1- 24 \ARMNUM d \exp\left(-\frac{1.8 s x^2}{16 \sigma^2} \right)$,
the following holds for all arms $i$ and attributes $j$
simultaneously:
$|\ArmE{\tau}{i}{j} - \Arm{i}{j}| \leq x$.
\dkcomment{We should clarify what the probability is over. It seems to
  be over the entire past history, but not the current type.}
\end{lemma}

@@

\begin{extraproof}{Lemma~\ref{lem:round-prob}}
Fix an arm $i$ and attribute $j$.
For any $n \geq 1$, let $k_n$ be the time step right after arm $i$ has
been pulled for the \Kth{n} time.
\dkcomment{I'm writing $n$ instead of $s$ here because $S_s$ would
  look ugly, and $n$ is consistent with the lemma.
  Or we could rename $S$ to something else.}
Define
$S_n := \frac{n \cdot (\ArmE{k_n}{i}{j} - \Arm{i}{j})}{2 \sigma}$
to be the sum of all attribute-$j$ noise components up to and
including the \Kth{n} pull of arm $i$,
\dkcomment{I'm not sure if we have defined the notation for components
  of vectors in the \emph{superscript}. Let's make sure not to forget it.}
renormalized to have standard deviation $1/2$.
\dkcomment{Is this the correct scaling.}
The $S_n$ define an unbiased half-subgaussian random walk,
and we can therefore apply Lemma~\ref{lem:ACI-inequality} to them.
Specifically, we set $b(n) = \frac{n x^2}{16 \sigma^2}$,
and obtain that

@@

Finally, this choice of $J$ ensures that
$\ArmE{k_J}{i}{j} = \ArmE{\tau}{i}{j}$,
and we have thus shown that
$|\ArmE{\tau}{i}{j} - \Arm{i}{j}| \leq x$ for all $i,j$.
The lemma now follows by a union bound over all \ARMNUM arms and $d$
attributes.
\end{extraproof}

Lemma~\ref{lem:round-prob} implies that errors from myopic play
decrease exponentially in the phase number $s$.

\begin{corollary} \label{cor:right-choice}
Let $x > 0$ be arbitrary.
Let $\LatePhase \geq \max (50, \frac{92.16 \sigma^4}{x^4})$.

Let $\tau$ be any stopping time that is almost surely \dkcomment{What
  exactly do we mean by this here?} in phase $s$,
with respect to the filtration
$\mathcal{F}_{t} = \sigma (\Pull{1}, \ldots, \Pull{t},
                          \PayA{1}, \ldots, \PayA{t},
                          \ObsV{1}, \ldots, \ObsV{t})$.
Then, for all $s \geq \LatePhase$,
with probability (over the past history) at least 
$1 - 24 \ARMNUM d \exp\left(-\frac{1.8 s x^2}{16 \sigma^2} \right)$,
\dkcomment{We should clarify what the probability is over. It seems to
  be over the past history, but not the current agent.}
all agent types $\AgV = \AgentV{\tau}$ with
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}}) > 2\Diam d x$
will pull their best arm, i.e.,
$\argmax \SET{\AgV \cdot \ArmEV{\tau}{i}} = \Best{\AgV}$
\end{corollary}

\begin{proof}
Let $\tau$ be a (random) time.
By Lemma~\ref{lem:round-prob},
with probability at least
$1 - 24 \ARMNUM d \cdot \exp \left( \frac{-1.8 s x^2}{16\sigma^2} \right)$,
for all arms $i$ and attributes $j = 1, \ldots, d$ simultaneously,
the estimated attribute $j$ for arm $i$ is close enough to the true attribute: 
$|\ArmE{\tau}{i}{j} - \Arm{i}{j}| \leq x$.
Condition on this high-probability event.

Consider any agent type $\AgV = \AgentV{\tau}$ at the (random) time
$\tau$ such that
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}}) > 2\Diam d x$.
Let $i \neq \Best{\AgV}$ be any other arm;
its true reward is much smaller than from the best arm:
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{i}) > 2\Diam d x$.
Because each coordinate of \ArmEV{\tau}{\Best{\AgV}} and of
\ArmEV{\tau}{i} is estimated accurately to within $x$, 
we get that 
$\AgV \cdot (\ArmEV{\tau}{\Best{\AgV}} - \ArmV{\Best{\AgV}})
\geq - \Diam d x$
and
$\AgV \cdot (\ArmEV{\tau}{i} - \ArmV{i}) \leq \Diam d x$.
Hence,

\begin{align*}
\AgV \cdot (\ArmEV{\tau}{\Best{\AgV}} - \ArmEV{\tau}{i})
& =
\AgV \cdot (\ArmEV{\tau}{\Best{\AgV}} - \ArmV{\Best{\AgV}})
+ \AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{i})
+ \AgV \cdot (\ArmV{i} - \ArmEV{\tau}{i})\\
& > -\Diam d x + 2 \Diam d x - \Diam d x
\; = \; 0,
\end{align*}
which means that the agent with type \AgV will in fact choose
the optimal arm in the high-probability event.
\end{proof}

@@

Towards bounding the algorithm's total payment, we first bound the
\emph{number} of rounds in which the algorithm makes a payment by a
constant.
This bound also turns out to be useful for bounding the total regret.

\begin{lemma} \label{lem:numP}
The expected number of time steps in which
Algorithm~\ref{alg:basic-incentivizing}
makes any payment is at most $O(\ARMNUM^2)$.
\end{lemma}

\begin{proof}
We partition phases into early and late phases.
For each of the early phases,
we crudely bound the number of payments by \ARMNUM,
using that each arm is incentivized at most once per phase.
For later phases,
we use that arms are payment-eligible only if their myopic pulling
probability is less than $1/s$.
Since at least a \MinProb fraction of agents would prefer arm $i$ if
given accurate estimates, this would require misestimates that make a
significant fraction of agents pull the wrong arm.
But in later phases, such misestimates are unlikely by
Corollary~\ref{cor:right-choice}.

To make this intuition precise, we set
$\delta = \frac{\MinProb}{2 \TieDensity}$,
and $x = \frac{\delta}{2 \Diam d}$.
The cutoff point between early and late phases is now set to
$\EvenLaterPhase = \max(50, \frac{92.16 \sigma^4}{x^4}, \frac{2}{\MinProb})$.
Consider the \Kth{s} pull of an arm $i$, for $s \geq \EvenLaterPhase$.
If this pull occurs before phase $s$, it was definitely not incentivized.
So we condition on the pull occurring in phase $s$,
and let $\tau$ be the time step when this pull occurred; notice that
this means that this must be the first pull of arm $i$ in phase $s$.
We can now apply Corollary~\ref{cor:right-choice}
with this choice of $x$,
and obtain that with probability at least 
$1 - 24 \ARMNUM d \cdot \exp\left(-\frac{1.8 s x^2}{16 \sigma^2} \right)$,
all agent types \AgV with
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}})
> 2\Diam d x = \delta$
will pull their best arm \Best{\AgV}.
\dkcomment{We are losing a factor \ARMNUM here.
We already took a union bound over all arms in the lemma, and
we now apply it only to one arm, then sum over all arms.}

By Lemma~\ref{lem:sdelta}, the \emph{total} measure of agents (across
all arms) whose best and second-best arm differ in utility by less
than $\delta$ is at most $\TieDensity \delta = \frac{\MinProb}{2}$.
In particular, this bound holds for agents whose best arm is $i$.
By Assumption~\ref{A3}, at least a measure \MinProb of agents has $i$
as their best arm, and thus, at least a measure $\frac{\MinProb}{2}$
has arm $i$ as their preferred arm by at least $\delta$ over any other
arm $i'$.
In the high-probability event, all such agents would myopically pull
the arm $i$, and as a result, in the high-probability event,
arm $i$ is not payment-eligible.

Thus, for any phase $s \geq \EvenLaterPhase$
Algorithm~\ref{alg:basic-incentivizing} incentivizes any given arm $i$
with probability at most
$24 \ARMNUM d \cdot \exp \left(
  -\frac{1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
\right)$.
Summing over all arms and phases $s$,
and adding the at most $\ARMNUM \EvenLaterPhase$ incentivizations in
the first \EvenLaterPhase phases, 
the expected total number of arm incentivizations is at most
\begin{align*}
\ARMNUM \EvenLaterPhase
  + 24 \ARMNUM^2 d \cdot \frac{1}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)}
& = O\left( \frac{\ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}
  + \frac{\ARMNUM^2 d}{1 - \exp \left(
    \frac{-1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)} \right).
\end{align*}
\dkcomment{It seems like with our choice of $x$ (which may be
  unavoidable), we get quite large factors, in particular with the
  fourth power.
  Also, it would be nice if we could simplify the $\frac{1}{1-\exp()}$
  term. Unfortunately, I think that the inequality goes the wrong way
  for applying the standard $\exp(-x) \geq 1-x$ bound.}
\end{proof}

It would be desirable to simply identify a constant upper bound on the
payment made each time.
Unfortunately, while the agent types are drawn from a bounded support,
the noise in arm locations is not;
hence, with small probability, arm locations may be grossly
misestimated, resulting in high incentive payments.
As a result, the actual analysis of the total payment is significantly more
intricate, and hence,
the proof of Theorem~\ref{rst:budget} is given in the appendix.

\subsection{Bounding the Total Regret}
\dkcomment{Edited up to here so far.}

### 2018-02-05 problem_setting.tex
finished edits up to payment

To support the formulation of our results and the analysis,
we define the following additional notation.
We let
$\Best{\AgentV{}} \in \argmax_i \AgentV{} \cdot \ArmV{i}$
and
$ \Second{\AgentV{}} \in \argmax_{i \neq \Best{\AgentV{}}} \AgentV{} \cdot \ArmV{i}$
denote the (indices of) the best and second-best arms for an agent
with attribute vector \AgentV{};
\dkdelete{breaking ties uniformly at random}.
\dkcomment{What would it mean to break ties at random in a
  definition?}

@@

Let $\MinProb = \min_{i} \Prob[\AgentDist]{\Set{\AgentV{}}{\Best{\AgentV{}} = i}}$
denote the minimum (over all arms) fraction of users that prefer any
particular arm.

@@

Let 
\begin{align*}
\FirstTwo{i}{i'}
  & = \Set{\AgentV{}}{\Best{\AgentV{}} = i, \Second{\AgentV{}} = i'}
\end{align*}
be the set of agent attribute vectors whose best arm is $i$ and
second-best arm is $i'$.
Let $F_{i,i'}$ be the marginal cumulative density function
(or cumulative mass function if $\AgentDist(\cdot)$ is a discrete distribution)
of $(\ArmV{i}-\ArmV{i'}) \cdot AgentV{}$,
conditioned on $\theta \in \Omega_{i,j}$.
In words, $F_{i,i'}$ is the distribution of the \emph{strength} of the
preference of a random agent for arm $i$ over arm $i'$, given that
$i$ and $i'$ are the agent's first and second choices.

\begin{assumption}[Not too many near-ties] \label{A1}
Near-ties have vanishing probability; 
that is, there exists a constant \TieDensity such that
for all pairs $i,i$ of arms and all $z \in \R^+$,
we have $F_{i,i'}(z) \leq \TieDensity \cdot z$.
\end{assumption}

>>>>

To support the formulation of our results and the analysis,
we define the following additional notation.
We let
$\Best{\AgV} \in \argmax_i \AgV \cdot \ArmV{i}$
and
$ \Second{\AgV} \in \argmax_{i \neq \Best{\AgV}} \AgV \cdot \ArmV{i}$
denote the (indices of) the best and second-best arms for an agent
with attribute vector \AgV;
\dkdelete{breaking ties uniformly at random}.
\dkcomment{What would it mean to break ties at random in a
  definition?}

@@

Let $\MinProb = \min_{i} \Prob[\AgentDist]{\Set{\AgV}{\Best{\AgV} = i}}$
denote the minimum (over all arms) fraction of users that prefer any
particular arm.

@@

Let $\FirstTwo{i}{i'} = \Set{\AgV}{\Best{\AgV} = i, \Second{\AgV} = i'}$
be the set of agent attribute vectors whose best arm is $i$ and
second-best arm is $i'$.
Let $F_{i,i'}$ be the marginal cumulative density function
(or cumulative mass function if $\AgentDist(\cdot)$ is a discrete distribution)
of $(\ArmV{i}-\ArmV{i'}) \cdot \AgV$,
conditioned on $\AgV \in \Omega_{i,j}$.
In words, $F_{i,i'}$ is the distribution of the \emph{strength} of the
preference of a random agent for arm $i$ over arm $i'$, given that
$i$ and $i'$ are the agent's first and second choices.

\begin{assumption}[Not too many near-ties] \label{A1}
Near-ties have vanishing probability; 
that is, there exists a constant \TieDensity such that
for all pairs $i,i'$ of arms and all $z \in \R^+$,
we have $F_{i,i'}(z) \leq \TieDensity \cdot z$.
\end{assumption}
